Factor Analysis on Citations, Using a Combined Latent and Logistic Regression Model

### 2019-08-22 Manuscript for submission/manuscript.tex
Update

\begin{remark}
Throughout the Section, we add $*$ to the superscript of parameters when they govern a data-generating process.
$\alpha^*$ is an intercept term in logistic regression model.
$f^*_j$ is the $j$th column of factor loading matrix $F^* \in \mathbb{R}^{K \times n}$, where $K$ refers to the number of topics embedded in network, and $n$ denotes the number of papers in a network.
Matrix $D^* \in \mathbb{R}^{K \times K}$ is a diagonal matrix whose entries denote weight coefficient of factors. 
$S_{ij}^*$ denotes an ad-hoc citational relation between $i$th and $j$th paper.
We use the notation $|S^*|$ to denote the cardinality of non-zero components of upper-triangular part of $S^*$ matrix 
(i.e., $|S^*|=|\{(i,j) : S_{ij}^{*}\neq 0, 1 \leq i<j \leq n\}|$).
We will use $X \sim \text{Unif}[a,b]$ to denote
$X$ is drawn from uniform distribution supported on the interval $[a,b]$.
\end{remark}

@@

\subsubsection{Synthetic Setting}
We consider following three scenarios: $\{(n^{(i)},K^{(i)},nnz^{(i)})\}_{i=1}^{3}=\{(30,3,10),(80,4,20),(120,5,40)\}$.

@@

One might wonder how traditional model selection methods, such as the Bayes Information Criterion (BIC;\cite{schwarz1978estimating}) and the AIC, work where we recall that BIC and AIC are defined as follows:
\[
\mbox{BIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + |M| \log \bigg(\frac{n(n-1)}{2}\bigg),
\]
and
\[
\mbox{AIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + 2|M|,
\]
where $M$ is the current model,
$\mathbb{L}_n( \hat{\beta}(M))$ is the maximal log-likelihood for a given model $M$,
and $|M|$ is the number of free parameters in $M$, which is determined by the number of non-zeros in $S$ and the low-rank matrix $L$.
If we have $\text{rank}(L)=K$, we can establish the following
\[
|M| = \sum_{i < j}1_{\{S_{ij} > 0\}} + n K - \frac{K(K-1)}{2} + 1 ;
\]
since the number of free parameters in $L$ is $K$ plus $nk - K(K+1)/2$, which is the number of free parameters in determining $K$ orth-normal vectors.
Additionally $1$ in the last is due to $\alpha$.
We want to find an $M$, which minimizes BIC($M$) or AIC($M$) as a function of $M$.
However these traditional methods seem to entail a serious problem.
%The assumptions used to derive BIC or AIC are not satisfied here, since each edges in a graph are not identically distributed.
Our experience shows that both BIC and AIC choose the most parsimonious model with the smallest $K(=\text{rank}(\widehat{L}))$ and the smallest cardinality$(=|\text{supp}(\widehat{S})|/2)$ of ad-hoc dependency among all the estimated ($\widehat{L}$,$\widehat{S}$)s'.
We evaluate the models that are selected via our heuristic approach, BIC, and AIC by using following three evaluation metrics.

>>>>

\begin{remark}
Throughout the Section, we add $*$ to the superscript of parameters when they govern a data-generating process.
$\alpha^*$ is an intercept term in logistic regression model.
$f^*_j$ is the $j$th column of factor loading matrix $F^* \in \mathbb{R}^{K \times n}$, where $K$ refers to the number of topics embedded in network, and $n$ denotes the number of papers in a network.
Here, we introduce another notation, $n_{l}$, denoting the number of papers, which share $l$ common topics.
Matrix $D^* \in \mathbb{R}^{K \times K}$ is a diagonal matrix whose entries denote weight coefficient of factors.
$S_{ij}^*$ denotes an ad-hoc citational relation between $i$th and $j$th paper.
We use the notation $|S^*|$ to denote the cardinality of non-zero components of upper-triangular part of $S^*$ matrix
(i.e., $|S^*|=|\{(i,j) : S_{ij}^{*}\neq 0, 1 \leq i<j \leq n\}|$).
We will use $X \sim \text{Unif}[a,b]$ to denote
$X$ is drawn from uniform distribution supported on the interval $[a,b]$.
\end{remark}

@@

\subsubsection{Synthetic Setting}
In first scenario, we consider following three cases: $\{(n_{1}^{(i)},K^{(i)},|S^*|^{(i)})\}_{i=1}^{3}$ $=\{(30,3,10),
(80,4,20),(120,5,40)\}$.
For example,  $(n_{1}^{(3)},K^{(3)},|S^*|^{(3)})$ means that we generate a network with $120$ papers, $5$ topics embedded, and $40$ arbitrary indexed ad-hoc links. Note $n_{1}^{(3)}=120$ means that each paper in the network has only one topic.
In second scenario, we consider a case where we have 
$(n,n_{2},n_{3},K,|S^{*}|)=(200,70,7,3,20)$.
Here, we generate a network with $200$ papers in total, which is comprised of $70$ papers with $2$ topics, $7$ papers with $3$ topics. The network has $3$ topics, and $20$ arbitrary indexed ad-hoc links. 
In third scenario, we consider a case:
$(n,n_{3},K,|S^{*}|)=(200,20,3,20)$.
In this scenario, we have a network with $200$ papers with $3$ topics. Among $200$ papers, there are $20$ papers, which are related with all the $3$ topics.
These $20$ papers have citational relationships with three clusters of papers via $20$ ad-hoc links, where papers in each cluster deals with only $1$ topic.

@@

One might wonder how traditional model selection methods, such as the Bayes Information Criterion (BIC;\cite{schwarz1978estimating}) and the AIC, work where we recall that BIC and AIC are defined as follows:
\[
\mbox{BIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + |M| \log \bigg(\frac{n(n-1)}{2}\bigg),
\]
and
\[
\mbox{AIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + 2|M|,
\]
where $M$ is the current model,
$\mathbb{L}_n( \hat{\beta}(M))$ is the maximal log-likelihood for a given model $M$,
and $|M|$ is the number of free parameters in $M$, which is determined by the number of non-zeros in $S$ and the low-rank matrix $L$.
If we have $\text{rank}(L)=K$, we can establish the following
\[
|M| = \sum_{i < j}1_{\{S_{ij} > 0\}} + n K - \frac{K(K-1)}{2} + 1 ;
\]
since the number of free parameters in $L$ is $K$ plus $nk - K(K+1)/2$, which is the number of free parameters in determining $K$ orth-normal vectors.
Additionally $1$ in the last is due to $\alpha$.
We want to find an $M$, which minimizes BIC($M$) or AIC($M$) as a function of $M$.
However these traditional methods seem to entail a serious problem.
%The assumptions used to derive BIC or AIC are not satisfied here, since each edges in a graph are not identically distributed.
Our experience shows that both BIC and AIC choose the most parsimonious model with the smallest $K(=\text{rank}(\widehat{L}))$ and the smallest cardinality$(=|\widehat{S}|)$ of ad-hoc dependency among all the estimated ($\widehat{L}$,$\widehat{S}$)s'.
We evaluate the models that are selected via our heuristic approach, BIC, and AIC by using following three evaluation metrics.
